### 2018-09-23 Sections/quantum_genetic_medicine.tex

Doing so will require highly complex processing of massive amounts of genetic and drug information, requiring demanding computational resources. We foresee quantum computing as playing a central role in the processing pipeline of genetic drug development, some initial ideas for which we present here.

@@

Following its own technological Moore's Law\index{Moore's Law}, one can reasonably anticipate this process becoming sufficiently cheap that in the near future it will become economically viable for every individual to have their own personal genome fully sequenced and available for medical use.

@@

\begin{figure}[!htbp]
	\includegraphics[width=0.475\textwidth]{genetic_sequencing_de_novo}
	\captionspacefig \caption{Three short-reads, each of 10 base-pairs in length (blue), which we wish to assemble into the largest possible contig\index{Contig} belonging to the underlying genome. The specificity\index{Specificity} defines the number of endpoint base-pairs that must overlap for them to be considered a match (orange). In this case the specificity is 3, and 3 short-reads have been assembled into a single 24 base-pair contig. The greater the specificity the more accurate will be the matching, but the lower the probability of finding matches. A specificity too low will create false matches, whereas a specificity too high makes finding sufficiently overlapping short-reads unlikely. This makes optimum choice of specificity an important consideration, creating a direct tradeoff between reconstructed contig length and accuracy. This sequencing technique is best applied when there is no known reference genome for the data under analysis. For the technique to work effectively there must be a sufficient quantity of short-read data that with high probability overlaps between reads will exist. If the data is too sparse few overlaps are likely to be found. Thus, average reconstructed contig lengths grow with the amount of short-read data available to the engine.} \label{fig:gen_seq_de_novo}
\end{figure}

@@

The failing of \textit{de novo} sequencing is that the pool of short-read data must be sufficiently large that all the pieces in the jigsaw puzzle are present, such that there are no gaps between neighbouring pieces -- contig `islands'\index{Contig islands} will remain isolated from other contigs. However, resources are of course always finite, as too is the number of short-reads available for reconstruction.

The other approach, \textit{mapping}, is to use an existing genome as a reference\index{Reference genome}. Rather than piece short-reads together, we compare them against this reference to infer their relative locations in the genome. When doing so, we allow some error threshold\index{Error threshold} in the matching. Specifically, we require the Hamming distance\index{Hamming distance} between a short-read and an equally-long segment of the reference be below some threshold. The flexibility offered by this threshold accommodates for mutational differences between the reference and the short-reads\index{Genetic mutations}, which is ultimately what we wish to characterise.

@@

\begin{figure}[!htbp]
	\includegraphics[width=0.475\textwidth]{genetic_sequencing_mapping}
	\captionspacefig \caption{In genome mapping we begin with a known reference genome (blue) to act as a comparison for short-reads (orange). The short-reads are slided along the reference until a location with sufficient overlap is found, at which point we record a match. By allowing some degree of single nucleotide mismatches (green) between the reference and short-read, we are able to characterise the SNPs in the short-read data. This sequencing technique is appropriate when we have a known genome for the species to which the short-read data belongs, but we wish to identify minor genetic mutations between the samples. For example, mapping short-reads belonging to human cancer cells to a human reference genome characterises the genetic mutations associated with that particular cancer. The technique works with any amount of short-read data, even just a single read, provided the reference is known.} \label{fig:gen_seq_mapping}
\end{figure}

@@

The key goal is to properly simulate the interaction of candidate drug compounds\index{Drug compounds}, ligands\index{Ligands}, or functional groups\index{Functional groups} with biochemical processes at the molecular level. Classical techniques for such simulations include estimating electron densities using density functional theory\index{Electron density}\index{Density functional theory} (the potential energy surface)\index{Electron density} of compounds, from which useful properties such as bonding affinities\index{Bonding affinity} may be determined.

@@

To achieve this, let us construct a classical algorithm for exhaustively constructing and enumerating organic drug molecules or functional groups, up to some cutoff size\index{Cutoff size}. The length of this list grows exponentially with the cutoff size. Next we implement this algorithm unitarily (which is always possible, writing the classical circuit as a reversible one), and construct it as a quantum oracle\index{Oracles} such that every input state (which acts as a pointer\index{Pointers} to an element in the list) yields as output a qubit representation\index{Qubit molecule representation} of the corresponding drug compound. Specifically, the oracle implements a transformation of the form,
\begin{align}
\hat{U}_\mathrm{oracle}\ket{i}_\mathrm{pointer}\ket{0}_\mathrm{drug}^{\otimes n} \to \ket{i}_\mathrm{pointer}\ket{\psi_i}_\mathrm{drug}^{(n)},
\end{align}
where $i$ is the index (pointer) to the drug element in the enumeration, and $n$ is the number of qubits encoding the representation of the drug molecule\footnote{This may be implemented unitarily, which is easy to see via its preservation of orthonormality.}.

@@

Having constructed this oracle, we feed the output qubit molecule description into the Hamiltonian simulation algorithm\index{Quantum simulation}, with exponential speedup, yielding as output a `score' characterising some desired property of the interaction, such as bonding affinity\index{Bonding affinity}. This implements the simulate-and-score unitary,
\begin{align}
\hat{U}_\mathrm{sim} \ket{\psi_i}_\mathrm{drug}\ket{0}_\mathrm{score} \to \ket{\psi_i}_\mathrm{drug}\ket{s_i}_\mathrm{score}	
\end{align}

@@

\begin{figure}[!htbp]
\includegraphics[width=0.475\textwidth]{genetic_medicine_pipeline}
	\captionspacefig \caption{Pipeline for quantum-enhanced genetic drug development. A quantum search algorithm searches over algorithmically-generated drug compounds or functional groups, for each performing Hamiltonian simulation, to determine if its score is above a threshold. The quantum search subroutine yields a quadratic enhancement in the number of drugs that can be simulated in parallel, and the Hamiltonian simulation subroutine yields an exponential enhancement over classical simulation techniques.} \label{fig:genetic_med_pipe}
\end{figure}

@@

In very recent years, new technologies have emerged for performing highly targeted gene editing, allowing genetic manipulations to be performed on living organisms. Using, CRISPR\index{CRISPR} technology, for example, one can enable and disable entire functional genes, or perform even more intricate manipulations, right down to the single nucleotide level.

>>>>

Doing so will require highly complex processing of massive amounts of genetic and drug information, with demanding computational resource requirements. We foresee quantum computing as playing a central role in the processing pipeline of genetic drug development, some initial ideas for which we sketch here.

@@

Following its own technological Moore's Law\index{Moore's Law}, one can reasonably anticipate this process becoming sufficiently cheap that in the near future it will become economically viable (and desirable) for every individual to have their own personal genome fully sequenced and available for medical use.

@@

\begin{figure}[!htbp]
	\includegraphics[clip=true, width=0.475\textwidth]{genetic_sequencing_de_novo}
	\captionspacefig \caption{Three short-reads, each of 10 base-pairs in length (blue), which we wish to assemble into the largest possible contig\index{Contig} belonging to the underlying genome. The specificity\index{Specificity} defines the number of endpoint base-pairs that must overlap for them to be considered a match (orange). In this case the specificity is 3, and 3 short-reads have been assembled into a single 24 base-pair contig. The greater the specificity the more accurate will be the matching, but the lower the probability of finding matches. A specificity too low will create false matches, whereas a specificity too high makes finding sufficiently overlapping short-reads unlikely. This makes optimum choice of specificity an important consideration, creating a direct tradeoff between reconstructed contig length and accuracy. This sequencing technique is best applied when there is no known reference genome for the data under analysis. For the technique to work effectively there must be a sufficient quantity of short-read data that with high probability overlaps between reads will exist. If the data is too sparse few overlaps are likely to be found. Thus, average reconstructed contig lengths grow with the amount of short-read data available to the engine.} \label{fig:gen_seq_de_novo}
\end{figure}

@@

The failing of \textit{de novo} sequencing is that the pool of short-read data must be sufficiently large that all the pieces in the jigsaw puzzle are present, such that there are no gaps between neighbouring pieces, whereby contig `islands'\index{Contig islands} remain isolated from other contigs. However, resources are of course always finite, as too is the number of short-reads available for reconstruction.

The other approach, \textit{mapping}, is to use an existing genome as a reference\index{Reference genome}. Rather than piece short-reads together, we compare them against this reference to infer their relative locations in the genome. When doing so, we allow some error threshold\index{Error threshold} in the matching. Specifically, we require the Hamming distance\index{Hamming distance} between a short-read and an equally-long segment of the reference be below some threshold. The flexibility offered by this threshold accommodates for mutational differences between the reference and the short-reads\index{Genetic mutations} (i.e the SNPs), which is ultimately what we wish to characterise.

@@

\begin{figure}[!htbp]
	\includegraphics[clip=true, width=0.475\textwidth]{genetic_sequencing_mapping}
	\captionspacefig \caption{In genome mapping we begin with a known reference genome (blue) to act as a comparison for short-reads (orange). The short-reads are slid along the reference until a location with sufficient overlap is found, at which point we record a match. By allowing some degree of single nucleotide mismatches (green) between the reference and short-read, we are able to characterise the SNPs in the short-read data. This sequencing technique is appropriate when we have a known genome for the species to which the short-read data belongs, but we wish to identify minor genetic mutations between the samples. For example, mapping short-reads belonging to human cancer cells to a human reference genome characterises the genetic mutations associated with that particular cancer. The technique works with any amount of short-read data, even just a single read, provided the reference is known.} \label{fig:gen_seq_mapping}
\end{figure}

@@

The key goal is to properly simulate the interaction of candidate drug compounds\index{Drug compounds}, ligands\index{Ligands}, or functional groups\index{Functional groups} with biochemical processes at the molecular level. Classical techniques for such simulations include estimating electron densities using density functional theory\index{Electron density}\index{Density functional theory} (the potential energy surface)\index{Electron density} of compounds, from which useful properties such as bonding affinities\index{Bonding affinity} may be estimated.

@@

To achieve this, let us construct a classical algorithm for exhaustively constructing and enumerating organic drug molecules or functional groups, up to some cutoff size\index{Cutoff size}. The length of this list grows exponentially with the cutoff size. Next we implement this algorithm unitarily (which is always possible, writing the classical circuit as a reversible one), and construct it as a quantum oracle\index{Oracles} such that every input state (which acts as a pointer\index{Pointers} to an element in the list) yields as output a qubit representation\index{Qubit molecule representation} of the corresponding drug compound. Specifically, the oracle implements a transformation of the form,
\begin{align}
\hat{U}_\mathrm{oracle}\ket{i}_\mathrm{pointer}\ket{0}_\mathrm{drug}^{\otimes n} \to \ket{i}_\mathrm{pointer}\ket{\psi_i}_\mathrm{drug}^{(n)},
\end{align}
where $i$ is the index (pointer) to the drug element in the enumeration, and $n$ is the number of qubits encoding the representation of the drug molecule\footnote{This may be implemented unitarily, which is easy to see via its preservation of orthonormality.} $\ket{\psi_i}_\mathrm{drug}^{(n)}$.

@@

Having constructed this oracle, we feed the output qubit molecule description into the Hamiltonian simulation algorithm\index{Quantum simulation}, with exponential speedup, yielding as output a `score' characterising some desired property of the interaction, such as bonding affinity\index{Bonding affinity}. This implements the simulate-and-score operator\index{Simulate-and-score operator},
\begin{align}
\hat{U}_\mathrm{sim} \ket{\psi_i}_\mathrm{drug}\ket{0}_\mathrm{score} \to \ket{\psi_i}_\mathrm{drug}\ket{s_i}_\mathrm{score}	
\end{align}

@@

\begin{figure}[!htbp]
\includegraphics[clip=true, width=0.475\textwidth]{genetic_medicine_pipeline}
	\captionspacefig \caption{Pipeline for quantum-enhanced genetic drug development. A quantum search algorithm searches over algorithmically-generated drug compounds or functional groups, for each performing Hamiltonian simulation, to determine if its score is above a threshold. The quantum search subroutine yields a quadratic enhancement in the number of drugs that can be simulated in parallel, and the Hamiltonian simulation subroutine yields an exponential enhancement over classical simulation techniques.} \label{fig:genetic_med_pipe}
\end{figure}

@@

In very recent years, new technologies have emerged for performing highly targeted gene editing, allowing genetic manipulations to be performed on living organisms. Using CRISPR\index{CRISPR} technology, for example, one can enable and disable entire functional genes, or perform even more intricate manipulations, right down to the single nucleotide level.

### 2018-09-23 Sections/to_the_future_and_beyond.tex

In this part we highlight some of -- in the minds of the authors -- some of the most exciting and transformative applications for the future quantum internet and its monstrous unified quantum computational power. None of the technologies discussed here have presently been realised on any practical scale, but all have revolutionary potential to transform the global economy, how we interact with the world, and how we live our lives.

>>>>

\dropcap{I}{n} this part we highlight some of -- in the minds of the authors -- the most exciting and transformative applications for the future quantum internet and its monstrous unified quantum computational power. None of the technologies discussed here have presently been realised on any practical scale, but all have revolutionary potential to transform the global economy, how we interact with the world, and how we live our lives.
